---
title: "Clustering"
author: "wz2270 Wenxuan Zhang"
date: "January 5, 2015"
output: html_document
---

```{r}
wine<-read.csv("wine.csv")
colnames(wine)<-c(
     "Cultivar",
     "Alcohol","Malic acid",
 	"Ash",
	"Alcalinity of ash",  
 	"Magnesium",
	"Total phenols",
 	"Flavanoids",
 	"Nonflavanoid phenols",
 	"Proanthocyanins",
	"Color intensity",
 	"Hue",
 	"OD280/OD315 of diluted wines",
 	"Proline")
wineTrain<-wine[,which(names(wine)!="Cultivar")]
#remove the item with correlation
set.seed(36563)
wineK3<-kmeans(x=wineTrain,centers=3)
wineK3
#install.packages("useful")
library(useful)
plot(wineK3,data=wineTrain)
plot(wineK3,data=wine,class="Cultivar")
set.seed(36563)
wineK3N25<-kmeans(wineTrain,centers=3,nstart=25)
wineK3$size
wineK3N25$size
```
Selecting number of clusters are of vital importance to success of clustering.
Choose number by Hartigan Rule.
compare within cluster sum of square with k and that of k+1 clusters.
FitKmeans can help you doing that
```{r}
wineBest<-FitKMeans(wineTrain,max.clusters=20,nstart=25,seed=278613)
PlotHartigan(wineBest)
table(wine$Cultivar,wineK3N25$cluster)
plot(table(wine$Cultivar,wineK3N25$cluster),main="Confusion Matrix for Wine Clustering",xlab="Cultivar",ylab="Cluster")
```
Another idea of selecting clusters is Gap statistic,which compares the within-cluster dissimilarity for a clustering of data with that of bootstraped sample of data.
```{r}
#install.packages("cluster")
library(cluster)
theGap<-clusGap(wineTrain,FUNcluster=pam,K.max=20)
gapDF<-as.data.frame(theGap$Tab)
gapDF
#logW curve
ggplot(gapDF,aes(x=1:nrow(gapDF)))+geom_line(aes(y=logW),color="blue")+geom_point(aes(y=logW),color="blue")+geom_line(aes(y=E.logW),color="green")+geom_point(aes(y=E.logW),color="green")+labs(x="Number of Clusters")
#gap curve
ggplot(gapDF,aes(x=1:nrow(gapDF)))+geom_line(aes(y=gap),color="red")+geom_point(aes(y=gap),color="red")+geom_errorbar(aes(ymin=gap-SE.sim,ymax=gap+SE.sim),color="red")+labs(x="Number of Clusters",y="Gap")
```
Kmean does not work with categlorical data and not robust. Thus there are K-medoids, which use real data point ,say median as center.
```{r}
indicators<-c("BX.KLT.DINV.WD.GD.ZS","NY.GDP.DEFL.KD.ZG","NY.GDP.MKTP.CD","NY.GDP.MKTP.KD.ZG","NY.GDP.PCAP.CD","NY.GDP.PCAP.KD.ZG","TG.VAL.TOTL.GD.ZS")
#install.packages("WDI")
library(WDI)
wbInfo<-WDI(country="all",indicator=indicators,start=2011,end=2011,extra=TRUE)
#get rid of aggregated info
wbInfo<-wbInfo[wbInfo$region!="Aggregates",]
#get rid of countries where all indicators are NA
wbInfo<-wbInfo[which(rowSums(!is.na(wbInfo[,indicators]))>0),]
#get rid of rows where the iso is missing
wbInfo<-wbInfo[!is.na(wbInfo$iso2c),]
rownames(wbInfo)<-wbInfo$iso2c
wbInfo$region<-factor(wbInfo$region)
wbInfo$income<-factor(wbInfo$income)
wbInfo$lending<-factor(wbInfo$lending)
keep.cols<-which(!names(wbInfo) %in% c("iso2c","country","year","capital","iso3c"))
wbPam<-pam(x=wbInfo[,keep.cols],k=12,keep.diss=TRUE,keep.data=TRUE)
#show the medoids observation
wbPam$medoids
#make a silhouette plot
plot(wbPam,which.plots=2,main="")
```
Programming download files.
```{r}
#This file did not work. try to download from the following source.
download.file(url="http://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/cultural/ne_10m_admin_0_sovereignty.zip",destfile="/Users/leilei/Clustering/ne_10m_admin_0_sovereignty.zip",method="curl")
unzip(zipfile="/Users/leilei/Clustering/ne_10m_admin_0_sovereignty.zip",exdir="/Users/leilei/Clustering/maps")
#install.packages("maptools")
library(maptools)
world<-readShapeSpatial("/Users/leilei/Clustering/world/world_borders.shp")
head(world@data)
world@data[which(world@data$CNTRY_NAME=="Austria"),]
#The name method of this map is different from world bank data 
library(plyr)
world@data$FIPS_CNTRY<-as.character(revalue(world@data$FIPS_CNTRY,replace=c(AS="AU",VM="VN",BM="MM",SP="ES",PO="PT",IC="IL",SF="ZA",TU="TR",IZ="IQ",UK="GB",EI="IE",SU="SD",MA="MG",MO="MA",JA="JP",SW="SE",SN="SG")))
library(ggplot2)
#install.packages("rgeos")
world@data$id<-rownames(world@data)
gpclibPermitStatus()
gpclibPermit()
gpclibPermitStatus()
#install.packages("gpclib")
library(gpclib)
world.df<-fortify(world,region="id")
#Method to convert a generic R object into a data frame useful for plotting. Takes its name from the idea of fortifying the original data with model fit statistics, and vice versa.
#
head(world.df)
world.df<-join(world.df,world@data[,c("id","CNTRY_NAME","FIPS_CNTRY")],by="id")
#just like join in sql, while join here is left join by default
head(world.df)
clusterMembership<-data.frame(FIPS_CNTRY=names(wbPam$clustering),Cluster=wbPam$clustering,stringsAsFactors=FALSE)
head(clusterMembership)
world.df=join(world.df,clusterMembership,by="FIPS_CNTRY")
world.df$Cluster<-as.character(world.df$Cluster)
world.df$Cluster<-factor(world.df$Cluster,levels=1:12)
ggplot()+geom_polygon(data=world.df,aes(x=long,y=lat,group=group,fill=Cluster,color=Cluster))+
    labs(x=NULL,y=NULL)+coord_equal()+theme(panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks=element_blank(),panel.background=element_blank())

#element_blank()
#Theme element: blank. This theme element draws nothing, and assigns no space
ggplot()+geom_polygon(data=world.df,aes(x=long,y=lat,group=group,fill=Cluster,color=Cluster))+
    labs(x=NULL,y=NULL)+coord_equal()
```
Determine number of clusters by cluster infromation
```{r}
wbPam$clusinfo
```
Hierarchical Clustering
```{r}
wineH<-hclust(d=dist(wineTrain))
plot(wineH)
#hierachical clustering for categlorical data
keep.cols<-which(!names(wbInfo)%in%c("iso2c","country","year","capital","iso3c"))
wbDaisy<-daisy(x=wbInfo[,keep.cols])
#compute all pair wise dissimilarityes between obseravtions in data set.
#daisy(x, metric = c("euclidean", "manhattan", "gower"),
 #     stand = FALSE, type = list(), weights = rep.int(1, p))
#if mixed(categlorical and numerical), use gower
#“Gower's distance” is chosen by metric "gower" or automatically if some columns of x are not numeric. Also known as Gower's coefficient (1971), expressed as a dissimilarity, this implies that a particular standardisation will be applied to each variable, and the “distance” between two units is the sum of all the variable-specific distances, see the details section.
wbH<-hclust(wbDaisy)
plot(wbH)
wineH1<-hclust(dist(wineTrain),method="single")
wineH2<-hclust(dist(wineTrain),method="complete")
wineH3<-hclust(dist(wineTrain),method="average")
wineH4<-hclust(dist(wineTrain),method="centroid")
plot(wineH1,labels=FALSE,main="Single")
plot(wineH2,labels=FALSE,main="Complete")
plot(wineH3,labels=FALSE,main="Average")
plot(wineH4,labels=FALSE,main="Centroid")

#either set number of clusters or where to cut
plot(wineH)
rect.hclust(wineH,k=3,border="red")
rect.hclust(wineH,k=13,border="blue")
plot(wineH)
rect.hclust(wineH,h=200,border="red")
rect.hclust(wineH,h=800,border="blue")
#fastcluster might be a good option
```





