---
title: "Time-series"
author: "wz2270 Wenxuan Zhang"
date: "January 16, 2015"
output: html_document
---

Most of these code are came from R for Everyone Jared P. Lander.
```{r}
require(WDI)
gdp<-WDI(country=c("US","CA","GB","DE","CN","JP","SG","IL"),indicator=c("NY.GDP.PCAP.CD","NY.GDP.MKTP.CD"),start=1960,end=2011)
names(gdp)<-c("iso2c","Country","Year","PerCapGDP","GDP")
require(ggplot2)
require(scales)
ggplot(gdp,aes(Year,PerCapGDP,color=Country,linetype=Country))+geom_line()+scale_y_continuous(label=dollar)
require(useful)
ggplot(gdp,aes(Year,GDP,color=Country,linetype=Country))+geom_line()+scale_y_continuous(label=multiple_format(extra=dollar,multiple="M"))
us<-gdp$PerCapGDP[gdp$Country=="United States"]
#convert us into timeseries
us<-ts(us,start=min(gdp$Year),end=max(gdp$Year))
plot(us,ylab="Per Capita GDP",xlab="Year")
```

ACF, aslo named autocorelation function,for a sequence x1,x2,x3,...xn. the zero order ACF is variance while the k order variance is caculate in the following way:$\sum_{i=1}^{i=n-k}{(x_{i+k}-\bar{x})}{(x_i-\bar{x})}/(n-1)$, k order correlation = k order variance/0 order variance.

Example
```{r}
a=c(1:10)
a=ts(a,start=1990,end=1999)
p<-acf(a)
p
secondordervar<-sum((a[c(1:9)]-mean(a))*(a[2:10]-mean(a)))/9
corr2<-secondordervar/var(a)
```

?PACF of order 2 , the effect that is not explained by the first lag.

```{r}
acf(us)
pacf(us)
x<-c(1,4,8,2,6,6,5,3)
diff(x,differences=1)
diff(x,differences=2)
diff(x,lag=1)
diff(x,lag=2)
```
Here are some difference between difference option and lag. Doing difference means:$x_{i+1}-x_i$
while lag=2 means:$x_{i+2}-x_i$

How to decide the optimal number difference?
```{r}
require(forecast)
ndiffs(x=us)
plot(diff(us,2))
usBest<-auto.arima(x=us)
acf(usBest$residuals)
pacf(usBest$residuals)
coef(usBest)
predict(usBest,n.ahead=5,se.fit=TRUE)
theForecast<-forecast(object=usBest,h=5)
plot(theForecast)
```
？Theory of arima

Multvariate Time Series
```{r}
require(reshape2)
gdpCast<-dcast(Year~Country,data=gdp[,c("Country","Year","PerCapGDP")],value.var="PerCapGDP")
head(gdpCast)
gdpTS<-ts(data=gdpCast[,-1],start=min(gdpCast$Year),end=max(gdpCast$Year))
#build a plot and legend using base graphics
plot(gdpTS,plot.type="single",col=1:8)
legend("topleft",legend=colnames(gdpTS),ncol=2,lty=1,col=1:8,cex=.4)
gdpTS<-gdpTS[,which(colnames(gdpTS)!="Germany")]

```
The most common way of fittig a modelto multiple time seriess is to use a vector auotregerssive model. The equation for a VAR is
$X_t=\phi_1X_{t-1}+...+\phi_pX_{t-p}+Z_t$
${Z_t~WN(0,\sum)}$ is white noise.

```{r}
numDiffs<-ndiffs(gdpTS)
gdpDiffed<-diff(gdpTS,differences=numDiffs)
plot(gdpDiffed,plot.type="single",col=1:7)
legend("bottomleft",legend=colnames(gdpDiffed),ncol=2,lty=1,col=1:7,cex=0.4)
#vars
#install.packages("vars")
require(vars)
gdpVar<-VAR(gdpDiffed,lag.max=12)
#how to define this lag.max
gdpVar$p
names(gdpVar$varresult)
class(gdpVar$varresult$Canada)
class(gdpVar$varresult$Japan)
head(coef(gdpVar$varresult$Canada))
head(coef(gdpVar$varresult$Japan))
#install.packages("coefplot")
require(coefplot)
coefplot(gdpVar$varresult$Canada)
coefplot(gdpVar$varresult$Japan)
predict(gdpVar,n.ahead=5)
```

GARCH
Generalized autoregressive conditional heteroskedasticity or GARCH family of models.
The model for variance in GARCH(m,s)

$/epsilon_t=/sigma_t*e_t$
where 
$/sigma_t^2=a_0+a_1/epsilon_{t-1}^2+...a_m\epsilon_{t-m}^2+\beta_1\sigma_{t-1}^2+...\beta_1\sigma_{t-1}^2+...\beta_s\sigma_{t-s}^2$
and
$e~GWN(0,1)$

```{r}
#install.packages("quantmod")
require(quantmod)
att<-getSymbols("T",auto.assign=FALSE)
require(xts)
head(att)
plot(att)
chartSeries(att)
addBBands()
addMACD(32,50,12)
attClose<-att$T.Close
class(attClose)
head(attClose)
install.packages("rugarch")
require(rugarch)
attSpec<-ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(1,1)),distribution.model="std")
attGarch<-ugarchfit(spec=attSpec,data=attClose)
plot(attGarch@fit$residuals,type='l')
plot(attGarch,which=10)
#ARMA
attSpec1<-ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(1,1)),distribution.model="std")
#ARMA(0.0)
attSpec2<-ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(0,0)),distribution.model="std")
#ARMA(0,2)
attSpec3<-ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(0,2)),distribution.model="std")
#ARMA(1,2)
attSpec4<-ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(1,2)),distribution.model="std")

attGarch1<-ugarchfit(spec=attSpec1,data=attClose)
attGarch2<-ugarchfit(spec=attSpec2,data=attClose)
attGarch3<-ugarchfit(spec=attSpec3,data=attClose)
attGarch4<-ugarchfit(spec=attSpec4,data=attClose)

infocriteria(attGarch1)
infocriteria(attGarch2)
infocriteria(attGarch3)
infocriteria(attGarch4)

attPred<-ugarchboot(attGarch,n.ahead=50,method=c("Partical","Full")[1])
#plot(attPred,which=2)

attLog<-diff(log(attClose))[-1]
attLogSpec<-ugarchspec(variance.model=list(model="sGARCH",garchOrder=c(1,1)),mean.model=list(armaOrder=c(1,1)),distribution.model="std")
#fit it
attLogGarch<-ugarchfit(spec=attLogSpec,data=attLog)
infocriteria(attLogGarch)

```
